Edge sign prediction in general graphs and Correlation Clustering

### 2017-12-02 cc/gtx_experiments.tex
reorganize files and write some intro
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

In this \namecref{sub:gtx_empirical_evaluation}, we provide empirical evidences of the properties of
\gtx{} over several classes of graph, and compare it with a \bfs{} baseline.\marginpars{If time
allows, it would be interesting to implement some methods of \vref{sub:gtx_state_of_the_art} and add
them to the comparison}\todo*{implement more low stretch methods} Namely, we consider three kinds of
graph topology (with both synthetic and real world instances that carry signs on their edges) and
evaluate $(i)$ what average stretch is reached by various trees and $(ii)$ how accurate is the sign
prediction.

@@

The three kinds of topology we consider are:
\begin{description}
	\item[\grid{}] which are 2D lattices, where each node has four neighbors except on the boundary.
		The synthetic ones are square, while the \enquote{real world} ones represents the four neighbors
		pixel connectivity of the pictures showed in \autoref{fig:gtx_xp_bwpics}.
	\item[\lpa{}] which are built synthetically according to the model of \textcite{Barabasi1999}.
		While this does not follow the more rigorous specification of \textcite{PAmodel04}, informally,
		we start with a line graph of $m$ nodes and add node one by one until the graph consists of $n$
		nodes. Each time a new node is added, it is connected to $m$ of the existing nodes with a
		probability proportional to their degree. Here we choose $m=3.13$, that is when adding a new
		node, we pick $3$ or $4$ existing neighbors such the initial expected number of neighbors for
		each new nodes is $3.13$. Such graphs are quite sparse and have short diameter, thus providing a
		crude but reasonable approximation of online social networks. Therefore, the real world
		instances of the \lpa{} model are \wik{}, \sla{} and \epi{}\marginpars{as used in the first
		chapter}\todo*{add a ref to first chapter} along with \gplus{}. The last one is constructed from
		ego networks of \gplus{}\footnote{Available at
		\url{http://snap.stanford.edu/data/egonets-Gplus.html}} by keeping the largest connected
		component of users whose gender is known. Basic statistics of those real \lpa{} graphs are
		presented in (\autoref{tab:gtx_xp_dataset}). 
	\item[\triangle{}] which consists of a Delaunay triangulation of random 2D points\footnote{As
		implemented by the \textsf{graph-tool} library (\url{https://graph-tool.skewed.de})}.
\end{description}

@@

The first property of Galaxy trees we wish to evaluate is their stretch, which depends only of graph
topology. Namely, let $G$ be a graph over vertex set $V$ with $|V|=n$ and edge set
$E$.\Todo[MOVE]{Stretch definition is likely to happen somewhere earlier} Furthermore, let $T$ be a
spanning tree of $G$ and $\etest{}$ the edges of $G$ not in $T$. Then we define the \emph{average
test edge stretch} as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}} |\mathrm{path}^T_{u,v}|$,
where $|\mathrm{path}^T_{u,v}|$ is the unique path between $u$ and $v$ in $T$.

As we consider unweighted graphs, we compare \gtx{} with a natural baseline, namely a spanning tree
rooted at the highest degree node and obtained through a breadth first visit of the graph. This
involves randomness in order in which nodes are visited. Likewise in \gtx{}, the choice of the edge
linking two stars is not always unique, meaning that we have to break ties at random.  Therefore,
for each graph, we repeat the tree construction 12 times and present the average result, noting that
the variance (showed as error bar in \autoref{fig:gtx_xp_st}) is small.

On \lpa{} and \triangle{}, we see that both trees exhibits logarithmic stretch, although with a
larger constant for \gtx{}. Note that this is also the case for others low stretch tree methods
\autocite[\S 5.3.1]{papplow}. On \grid{} however, \gtx{} preserves this logarithmic stretch growth
while this is visually no longer the case for \bfs{}.
In that case, we cannot expect a better stretch than $\frac{\log n}{2048}$ according to
\autocite[Theorem 6.6]{LowerBound95}.

@@

The second design goal of Galaxy trees is to accurately predict the sign of edges in $\etest{}$.
Except for the three real datasets that already include signs\footnote{We nonetheless perform some
preprocessing in order to make them undirected to remove the small proportion of conflicting edges
(e.g. positive from $u$ to $v$ but negative from $v$ to $u$).}, all the other are constructed,
meaning we have to set sign on their edges in the first place. This is done by partitioning the
nodes into two clusters. For \gplus{} we use node gender, for pictures we use node color (black or
white), and for all others, we propagate labels $0$ and $1$ from randomly selected high degree nodes.
Once each node belongs to one of the two clusters, we set the sign of an edge between two nodes to
be $+$ if they are in the same cluster and $-$ otherwise.  Predicting using path parity will thus
gives perfect result. To test performance in real or adversarial situation, we then add noise, that
is we select a fraction of edges uniformly at random and flip their sign.

We evaluate the performance of our prediction using the Matthews Correlation\Todo{merge this
definition of MCC with the one in first chapter} Coefficient (MCC)~\autocite{MCC00} \[
	\mathrm{MCC} = \frac{ TP \times TN - FP \times FN } {\sqrt{ (TP + FP) ( TP + FN ) (
			TN + FP ) ( TN + FN ) } } = \pm \sqrt{\frac{\chi^2}{n}}
\]
Since we do not have confidence score, we cannot use AUC. Yet we have to account for the large sign
unbalance and thus cannot rely on accuracy or $F_1$ measure.  Therefore we choose MCC, which
combines all the four numbers of the confusion matrix in a single metric. It ranges from $+1$
(perfect prediction) to $-1$ (inverse prediction) through $0$ (random prediction). As a
demonstration of MCC usefulness, predicting all edges but one to be positive on Slashdot gives
$.764$ accuracy, $.886$ $F_1$ score\marginpars{Actually the $F_1$ score is $.866$ for positive
edges, $0$ for negative ones and $.661$ if we can take an average weighted by class size.} but
$-0.0007$ MCC.

As showed in \autoref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
\bfs{}. As the noise level gets higher, they have similar performance. Note also than in
\autoref{fig:gtx_xp_pasynthmcc}, \gtx{} is less sensible to the size of the graph.

@@

To further assess the quality of our trees, we plug them in them into a successful heuristic method
to predict edge sign: \asym{}~\autocite{Kunegis2009}. \Todo{It might also be interesting to see if
that would be a good training set for our troll method, although it has to be checked it makes sense
from a running time point of view.} It computes the exponential of the adjacency matrix after it has
been reduce to $z$ dimension. This allows to count the sign of all paths between two pairs of nodes
with decreasing weight depending of their length. To simulate an active learning setting, we reveal
only a subset of edge in $A$. This subset can be: $i)$ the edges forming a \bfs{}, $ii)$ the edges
forming a \gtx{} $iii)$ $|V|-1$ edges chosen uniformly at random.

We set the parameter $z$ equal to $15$ because $i)$ it is one of the best in \cite[Fig.
11]{Kunegis2009}, $ii)$ it performs well on real dataset in \cite[Fig.3]{Cesa-Bianchi2012a}, and
$iii)$ it was good in our initial testing (\texttt{20150401\_wed\_spectral.ipynb}).

As the \asym{} has a $O(n^3)$ complexity and uses quite some memory at prediction time, the larger
graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} is that
except on social network, it is better to use spanning trees than random edges. Specifically, \gtx{}
on \grid{} and \bfs{} elsewhere.

@@

Finally\marginpars{Actually I never did it because \shz{} wasn't implemented at the time, so now is
a good occasion}\todo*{Run shazoo on galaxy tree} we also compare \gtx{} with \bfs{} and \rst{} on
the task of nodes prediction using \shz{} algorithm~\autocite{Vitale2012}.

>>>>

In this section, we provide empirical evidences of the properties of
\gtx{} over several classes of graph, and compare it with a \bfs{} baseline.
%\marginpars{If time allows, it would be interesting to implement some methods of \vref{sub:gtx_state_of_the_art} and add them to the comparison}\todo*{implement more low stretch methods}
Namely, we consider three kinds of
graph topology (with both synthetic and real world instances that carry signs on their edges) and
evaluate $(i)$ what average stretch is reached by various trees and $(ii)$ how accurate is the sign
prediction.

@@

The three kinds of topology we consider are:
\begin{description}[leftmargin=*]
	\item[\grid{}] which are 2D lattices, where each node has four neighbors except on the boundary.
		The synthetic ones are square, while the \enquote{real world} ones represents the four neighbors
		pixel connectivity of the pictures showed in \autoref{fig:gtx_xp_bwpics}.
	\item[\lpa{}] which are built synthetically according to the model of \textcite{Barabasi1999}.
		While this does not follow the more rigorous specification of \textcite{PAmodel04}, informally,
		we start with a line graph of $m$ nodes and add node one by one until the graph consists of $n$
		nodes. Each time a new node is added, it is connected to $m$ of the existing nodes with a
		probability proportional to their degree. Here we choose $m=3.13$, that is when adding a new
		node, we pick $3$ or $4$ existing neighbors such the initial expected number of neighbors for
		each new nodes is $3.13$. Such graphs are quite sparse and have short diameter, thus providing a
		crude but reasonable approximation of online social networks. Therefore, the real world
		instances of the \lpa{} model are \wik{}, \sla{} and \epi{} networks from \autoref{chap:troll}
		 along with \gplus{}. The last one is constructed from
		ego networks of \gplus{}\footnote{Available at
		\url{http://snap.stanford.edu/data/egonets-Gplus.html}} by keeping the largest connected
		component of users whose gender is known. Basic statistics of those real \lpa{} graphs are
		presented in (\autoref{tab:gtx_xp_dataset}). 
	\item[\triangle{}] which consists of a Delaunay triangulation of a set of points randomly
		located in a 2D space.\footnote{As
		implemented by the \textsf{graph-tool} library (\url{https://graph-tool.skewed.de})}
\end{description}

@@

The first property of Galaxy trees we wish to evaluate is their stretch, which depends only of graph
topology. Recall that following equation~\autoref{eq:test_stretch_def}, we define the average test
edge stretch as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}} |\mathrm{path}^T_{u,v}|$,
where $|\mathrm{path}^T_{u,v}|$ is the unique path between $u$ and $v$ in $T$.

As we consider unweighted graphs, we compare \gtx{} with a natural baseline, namely a spanning tree
rooted at the highest degree node and obtained through a breadth first visit of the graph. This
involves randomness in the order in which nodes are visited. Likewise in \gtx{}, the choice of the edge
linking two stars is not always unique, meaning that we have to break ties at random.  Therefore,
for each graph, we repeat the tree construction \np{12} times and present the average result, noting that
the variance (showed as error bar in \autoref{fig:gtx_xp_st}) is small.

On \lpa{} and \triangle{}, we see in \autoref{fig:gtx_xp_st} that both trees exhibits logarithmic stretch, although with a
larger constant for \gtx{}. Note that this is also the case for others low stretch tree methods
\autocite[Section 5.3.1]{papplow}. On \grid{} however, \gtx{} preserves this logarithmic stretch growth
while this is visually no longer the case for \bfs{}.
In that case, we cannot expect a better stretch than $\frac{\log n}{2048}$ according to
\autocite[Theorem 6.6]{LowerBound95}.

@@

The second design goal of Galaxy trees is to accurately predict the sign of edges in $\etest{}$.
Except for the three real datasets that already include signs\footnote{We nonetheless perform some
preprocessing in order to make them undirected and to remove the small proportion of conflicting edges
(e.g. positive from $u$ to $v$ but negative from $v$ to $u$).}, all the other are constructed,
meaning we have to set sign on their edges in the first place. This is done by partitioning the
nodes into two clusters. For \gplus{} we use node gender, for pictures we use node color (black or
white), and for all others, we propagate labels $0$ and $1$ from randomly selected high degree nodes.
Once each node belongs to one of the two clusters, we set the sign of an edge between two nodes to
be $+$ if they are in the same cluster and $-$ otherwise.  Predicting using path parity will thus
gives perfect result. To test performance in real situations, we then add noise, that
is we select a fraction of edges uniformly at random and flip their sign.

Like in \autoref{s:exp}, we evaluate the performance of our prediction using the Matthews
Correlation Coefficient (MCC), defined in equation~\autoref{eq:troll_mcc} \vpageref{eq:troll_mcc}.
As showed in \autoref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
\bfs{}. As the noise level gets higher, they have similar performance. Note also than in
\autoref{fig:gtx_xp_pasynthmcc}, \gtx{} is less sensible to the size of the graph.

@@

To further assess the quality of our trees, we plug them in them into an existing heuristic method
to predict edge sign: \asym{}~\autocite{Kunegis2009}.
%\Todo{It might also be interesting to see if that would be a good training set for our troll method, although it has to be checked it makes sense from a running time point of view.}
It computes the exponential of the adjacency matrix after it has
been reduce to $z$ dimension. This allows to count the sign of all paths between two pairs of nodes
with decreasing weight depending of their length. To simulate an active learning setting, we reveal
only a subset of edge in $A$. This subset can be: $i)$ the edges forming a \bfs{}, $ii)$ the edges
forming a \gtx{} $iii)$ $|V|-1$ edges chosen uniformly at random.
We set the parameter $z$ equal to $15$ because $i)$ it is one of the best in \cite[Fig.
11]{Kunegis2009} and $ii)$ it performs well on real datasets in \cite[Fig.3]{Cesa-Bianchi2012a}.
% , and $iii)$ it was good in our initial testing (\texttt{20150401\_wed\_spectral.ipynb}).

As the \asym{} has a $O(n^3)$ complexity and uses quite some memory at prediction time, the larger
graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} is that
except on social networks, it is better to use spanning trees than random edges. Specifically, \gtx{}
on \grid{} and \bfs{} elsewhere.

@@

\iffalse
Finally\marginpars{Actually I never did it because \shz{} wasn't implemented at the time, so now is
a good occasion}\todo*{Run shazoo on galaxy tree} we also compare \gtx{} with \bfs{} and \rst{} on
the task of nodes prediction using \shz{} algorithm~\autocite{Vitale2012}.
\fi

### 2017-12-02 cc/gtx_main.tex
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

Given a graph $G$, and following the social balance theory roughly summarized as \enquote{the enemy
of my enemy is my friend}, assume that the labelling of $E$ is consistent with a two-clustering of
$V$. Namely, assume that $V$ can be partitioned in two clusters such that edges within each cluster
are positive and edges across clusters are negative. In that case, the following
\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$ (we call this product the
parity of $p$ and denote it $\pi(p)$). While it is a simple and convenient model, this is too strong
of a requirement to be satisfied in practice. Therefore, we relax it by assuming that we start with
a consistent labeling $Y$ but can only observe a randomly perturbed version $Y'$ of $Y$.
Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, each sign of $Y$ is flipped with a
probability smaller than $q$.

In this section, we are interested in active learning algorithms, that first query a subset
$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
we describe a way to build spanning trees tailored for this situation.

>>>>

We now show how to apply the learning bias of \autoref{sec:new_bias} on undirected graphs.
Furthermore, we assume that the strong balance holds, meaning that there are only two groups
according to \autoref{thm:structural}. In other words, the labelling of $E$ is consistent with a
two-clustering of $V$. Namely, $V$ can be partitioned in two clusters such that edges within each
cluster are positive and edges across clusters are negative.
In that case, the following
\emph{multiplicative rule} holds: for any nodes $u$, $v$ in $V$, and any path $p$ between $u$ and
$v$ in $G$, the sign $\yuv$ is equal to the product of the signs along $p$. Hereafter, we call this
product the parity of $p$, and denote it by $\pi(p)$.
While it is a simple and convenient hypothesis, this is too strong
of a requirement to be satisfied in practice. Therefore, we relax it by assuming that, starting from
a consistent labeling $Y$, we can only observe a randomly perturbed version $Y'$ of $Y$.
Specifically, given a constant $q\in [0, \nicefrac{1}{2})$, every sign of $Y$ is flipped with a
probability smaller than $q$. We denote by $E_{\mathrm{flip}} \subset E$ the set of edges whose sign
has been flipped.

In this section, we are interested in active learning algorithms that first query a subset
$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
More precisely, we focus on an algorithm that queries a spanning tree $T$ of $G$ and predicts the
sign of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. Intuitively, since
each sign has been potentially flipped, the longer the path in $T$, the more likely its parity will
be not be equal to the true sign $\yuv{}$. Therefore we would like each such path to be as short as
possible.
Formally, the number of mistakes of such an algorithm is upper bounded by~\autocite[Equation
(3)]{Cesa-Bianchi2012b}
\begin{equation*}
  |E_{\mathrm{flip}}| + \sum_{(u,v) \in \etest}
  \sum_{e\in E} \Ind{e \in \pathtuv} \Ind{e \in E_{\mathrm{flip}}}
\end{equation*}
which in expectation is equal to:
\begin{equation}
  \label{eq:stretch_mistakes}
  q\left(|E| + \sum_{(u,v) \in \etest} |\pathtuv| \right)
\end{equation}
In the following, we describe a way to build spanning trees tailored for this situation. More
precisely, we implement and analyze a suggestion made to us by~\autocite{gtxFabio}.

### 2017-12-02 troll/experiments.tex
review gtx description and experiments
in particular, avoid mentioning strech by name, and say we only have a
conjecture regarding the runtime
also make the transition between cc and gtx smoother

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are
positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We
instead evaluated our performance using the so-called Matthews Correlation Coefficient
(MCC)~\autocite{MCC00}, defined as
\[
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }% = \pm \sqrt{\frac{\chi^2}{n}}
\]
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue
$n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$
(when all predictions are incorrect) to $+1$ (when all predictions are correct) through $0$ (when
predictions are made uniformly at random).

>>>>

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are
positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We
instead evaluated our performance using the so-called Matthews Correlation Coefficient
(MCC)~\autocite{MCC00}, defined as
\begin{equation}
  \label{eq:troll_mcc}
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }% = \pm \sqrt{\frac{\chi^2}{n}}
\end{equation}
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue
$n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$
(when all predictions are incorrect) to $+1$ (when all predictions are correct) through $0$ (when
predictions are made uniformly at random).
